Parametricity Consequences of GADTs Representations

### 2021-02-16 20-gadts/GADT-story.tex
setting up Par section for Enrico

@@

@@

2. Non-uniformity renders build, therefore short cut, for GADTs
impossible. Can't even formulate! It is possible there could be
another way to prove equivalence, but we know the two representations
are not equivalent because we have a counterexample.

The fact that GADTs don't validate any such iso has important
consequences for their parametricity properties.

>>>>

{\color{blue} THIS TRANSITION FROM EQUIVALENCE OF REPRESENTATION TO
  PARAMETRICITY GOES SOMEWHERE:
Non-uniformity renders build, therefore short cut, for GADTs
impossible. Can't even formulate! It is possible there could be
another way to prove equivalence, but we know the two representations
are not equivalent because we have a counterexample.

The fact that GADTs don't validate any such iso has important
consequences for their parametricity properties.}

@@

{\color{blue} Say somewhere: A relation is a triple $(A,B,R)$, where
  $R$ is a relation between sets $A$ and $B$.  A function from
  $(A,B,R)$ to $(A',B',R')$ is a pair $(f : A \to A',g : B \to B')$ of
  functions on sets such that $(f a,g\,b) \in R'$ if $(a,b) \in R$. We
  may write $R : \rel(A,B)$ for $(A,B,R)$.  If $R : \rel(A,B)$ we
  write $\pi_1 R$ and $\pi_2 R$ for the {\em domain} $A$ of $R$ and
  the {\em codomain} $B$ of $R$, respectively, and assume $\pi_1$ and
  $\pi_2$ are surjective. We write $\Eq_A = (A,A,\{(x,x)~|~ x \in
  A\})$ for the {\em equality relation} on the set $A$.}

The key idea underlying Reynolds' parametricity is to give each type
$F(\alpha)$ with one free variable $\alpha$ a {\em set interpretation}
$F_0$ taking sets to sets and a \emph{relational interpretation} $F_1$
taking relations $R : \rel(A,B)$ to relations $F_1 (R) : \rel(F_0 (A),
F_0 (B))$, and to interpret each term $t(\alpha,x) : F(\alpha)$ with
one free term variable $x : G(\alpha)$ as a map $t_0$ associating to
each set $A$ a function $t_0(A) : G_0(A) \to F_0(A)$. These
interpretations are given inductively on the structures of $F$ and $t$
in such a way that they imply two fundamental theorems. The first is
an \emph{Identity Extension Lemma}, which states that $F_1(\Eq_A) =
\Eq_{F_0(A)}$, and is the essential property that makes a model
relationally parametric rather than just induced by a logical
relation. The second is an \emph{Abstraction Theorem}, which states
that, for any $R :\rel(A, B)$, $(t_0(A),t_0(B))$ is a morphism in
$\rel$ from $(G_0(A),G_0(B),G_1(R))$ to $(F_0(A),F_0(B),F_1(R))$. The
Identity Extension Lemma is similar to the Abstraction Theorem except
that it holds for {\em all} elements of a type's interpretation, not
just those that interpret terms.  Similar theorems are required for
types and terms with any number of free variables.

@@

2. What are the parametricity properties for GADTs in each setting?
(Naturality only with primitives, inhabitation etc. with CEs)

{\color{red} ENRICO: Just write a section that can go here about
  1. for our favorite GADT.}

\vspace*{1in}

NOTES/POINTS:
